---
title: "Cold Spell at 10-days Post-infection"
output: html_document
date: '2023-08-04'
---

#Library Set Up

```{r Library Setup}
#creating percentages and tidying data
library(dplyr)
#glm.nb model
library(MASS)
#elastic regression
library(glmnet)
#AUC binomial model check
library(pROC)
#feature importance 
library(caret)
#post-hoc contrasts for models 
library(emmeans)
#zero-inflated model
library(pscl)
#GGPlot for graphing
library(ggplot2)
```

#Import and Clean Data 

````{r Import Data Set}
coldsnap<-read.csv("CS_Data.csv", 
                 header = TRUE, 
                 stringsAsFactors = FALSE)
```

```{r Remove early deaths, males and uninfected controls}
CST <- subset(coldsnap, Include == "Yes")
```

```{r Make Treatment Column}
CST$Treatment <- paste(CST$Amplitude, CST$Duration)
```

```{r Set Variables}
CST$A_Temp<-as.numeric(CST$A_Temp)
CST$Real_Avg<-as.numeric(CST$Real_Avg)
CST$Bath<-as.factor(CST$Bath)
CST$Timing<-as.factor(CST$Timing)
CST$Amplitude<-as.factor(CST$Amplitude)
CST$Duration<-as.factor(CST$Duration)
CST$Infection<-as.numeric(CST$Infection)
CST$Spores<-as.numeric(CST$Spores)
CST$Exposed<-as.numeric(CST$Exposed)
CST$Treatment <- as.factor(CST$Treatment)
```

```{r Specifying Order of Factors}
CST$Amplitude <- factor(CST$Amplitude, levels = c("0", "-3", "-6"))
CST$Duration <- factor(CST$Duration, levels = c("0", "3", "6"))
CST$Treatment <- factor(CST$Treatment, levels = c("0 0", "-3 3", "-3 6", "-6 3", "-6 6"))
```

```{r Adding Proportion Infected and Standard Error of Burden Column}
CST = CST %>%
  group_by(Treatment, A_Temp, Timing) %>% 
  #Make proportion infected columns 
  add_count(Treatment) %>% 
  mutate(
    nn = sum(Infection != 0),
    Prop = nn/n,
  #creating binomial confidence intervals
  se.I = sqrt(Prop * (1-Prop)/n),
  lower.I = pmax(Prop - (se.I * 1.96), 0),
  upper.I = pmin(Prop + (se.I * 1.96), 1)
  ) %>% 
  
  #Make standard error columns
  mutate(
    mean = mean(Spores, na.rm = T),
    sd.B = sd(Spores, na.rm = T),
    se.B = sd.B/sqrt(n())) %>%
  ungroup()
```

```{r Centre Average Temperature}
# Centering the 'Real_Avg' variable
CST$Real_Avg_Cent <- scale(CST$Real_Avg, center = TRUE, scale = FALSE)
CST$Real_Avg_Cent <- as.numeric(CST$Real_Avg_Cent)
```

```{r Extra Subsets}
#Subset for only core average temperatures
CSTC <- subset(CST, A_Temp %in% c("14","17","20","23"))
CSTC$A_Temp <- as.factor(CSTC$A_Temp)

# Remove NAs from the data set not relevant as we are only looking at infected individuals
CST.NA <- na.omit(CST)
```

#Check Baths

```{r Temperature and Bath Interactions for Infection}
Bath.CI <- glm(Infection ~ Bath, family = binomial, data = CST)
summary(Bath.CI)

Bath.Test.CI <- anova(Bath.CI, test = "Chi")
Bath.Test.CI
```

```{r Temperature and Bath Interactions for Burden}
Bath.CB <- glm.nb(Spores ~ Bath, data = CST)
summary(Bath.CB)

Bath.Test.CB <- anova(Bath.CB, test = "Chi")
Bath.Test.CB
```

#Infection Model

```{r Centred Infection Model}
#using a quadratic polynomial as it gives the lowest AIC out of the suitable models
ten.quadratic.CI <- glm(Infection ~ poly(Real_Avg_Cent, 2) * Amplitude * Duration, family = binomial, data = CST)
summary(ten.quadratic.CI)
```

From the basis of this model an elastic regression is created with high colinearity so need a new approach 

```{r Elastic Net Regression for Infection}
# Prepare the data
x.CI <- model.matrix(Infection ~ poly(Real_Avg_Cent, 2) * Amplitude * Duration, data = CST)  # Create the design matrix
y.CI <- CST$Infection  # Extract the response variable
```

```{r Finding the optimal aplha}
#trying various alphas to find which one is the most suitable 
foldid.CI<- sample(1:10, size = length(y.CI), replace = TRUE)
cv1.CI  <- cv.glmnet(x.CI, y.CI, foldid = foldid.CI, family = "binomial", alpha = 1)
cv.5.CI <- cv.glmnet(x.CI, y.CI, foldid = foldid.CI, family = "binomial", alpha = 0.5)
cv0.CI  <- cv.glmnet(x.CI, y.CI, foldid = foldid.CI, family = "binomial", alpha = 0)

par(mfrow = c(2,2))
plot(cv1.CI); plot(cv.5.CI); plot(cv0.CI)
plot(log(cv1.CI$lambda)   , cv1.CI$cvm , pch = 19, col = "red",
     xlab = "log(Lambda)", ylab = cv1.CI$name)
points(log(cv.5.CI$lambda), cv.5.CI$cvm, pch = 19, col = "grey")
points(log(cv0.CI$lambda) , cv0.CI$cvm , pch = 19, col = "blue")
legend("topleft", legend = c("alpha= 1", "alpha= .5", "alpha 0"),
       pch = 19, col = c("red","grey","blue"))
```

So alpha 0.5 deviates less and has the lowest log(lambda) so is the most suitable. 

```{r Model Fitting}
# Elastic regression model 
elfit.b.CI <- cv.glmnet(x.CI, y.CI, family = "binomial", alpha = 0.5)  # Fit a ridge regression model

# Select optimal tuning pattern (lambda) for ridge regression
opt_lambda.CI <- elfit.b.CI$lambda.min  # Optimal lambda value for regression via CV

# Refit the model using the selected lambda
elfit.fin.CI <- glmnet(x.CI, y.CI, family = "binomial", alpha = 0.5, lambda = opt_lambda.CI)  
# Fit the ridge regression model with the selected lambda
```

```{r bootstrapped significance testing for the coefficients of infection}
# Number of bootstrap iterations
num_bootstraps.CI <- 25000

# Empty list to store bootstrap coefficients
bootstrap_coefs_list.CI <- vector("list", length = num_bootstraps.CI)

# Set the seed for reproducibility
set.seed(123)

# Perform bootstrapping
for (i in 1:num_bootstraps.CI) {
  # Sample with replacement from the data
  sample_indices.CI <- sample(nrow(x.CI), replace = TRUE)
  x_sampled.CI <- x.CI[sample_indices.CI, ]
  y_sampled.CI <- y.CI[sample_indices.CI]
  
  # Fit the elastic net model on the sampled data
  model_fit.CI <- glmnet(x_sampled.CI, y_sampled.CI, family = "binomial", alpha = 0.5, lambda = opt_lambda.CI)
  
  # Store the coefficients in the list
  bootstrap_coefs_list.CI[[i]] <- as.vector(coef(model_fit.CI, s = opt_lambda.CI))
}

# Convert the list of bootstrap coefficients to a matrix
bootstrap_coefs.CI <- do.call(rbind, bootstrap_coefs_list.CI)

# Calculate the 95% confidence intervals for each coefficient
lower_ci.CI <- apply(bootstrap_coefs.CI, 2, function(coef_col) quantile(coef_col, 0.025))
upper_ci.CI <- apply(bootstrap_coefs.CI, 2, function(coef_col) quantile(coef_col, 0.975))

# Convert the coefficients to a matrix
coef_matrix.CI <- as.matrix(coef(elfit.fin.CI, s = opt_lambda.CI))

# Combine the results into a data frame for easy interpretation
coef_summary.CI <- data.frame(Coefficient = colnames(coef_matrix.CI),
                              Estimate = coef_matrix.CI,
                              Lower.CI = lower_ci.CI,
                              Upper.CI = upper_ci.CI)

# Print the summary
print(coef_summary.CI)
```

```{r Feature Importance for Infection}
# Extract the coefficients from the final model
coef_matrix.CI <- as.matrix(coef(elfit.fin.CI, s = opt_lambda.CI))

# Get the absolute values of coefficients to identify feature importance
feature_importance.CI <- abs(coef_matrix.CI[-1,])

# Sort the features by importance (highest to lowest)
sorted_features.CI <- feature_importance.CI[order(-feature_importance.CI)]

# Get the corresponding feature names
sorted_feature_names.CI <- colnames(coef_matrix.CI)[-1][order(-feature_importance.CI)]

# Combine the feature names with the sorted feature importance values
sorted_feature_importance_table.CI <- data.frame(Feature = sorted_feature_names.CI,
                                                 Importance = sorted_features.CI)

# Print the sorted feature importance table
print(sorted_feature_importance_table.CI)
```

```{r Model Fit (Area Under the Curve)}
# Calculate the cross-validated AUC score using the training data
# The predict() function with type = "response" will give cross-validated predicted probabilities
cv_predicted_probabilities.CI <- predict(ridge_model.CI, s = "lambda.min", newx = x.CI, type = "response")
cv_roc_curve.CI <- roc(y.CI, cv_predicted_probabilities.CI)
cv_auc_score.CI <- auc(cv_roc_curve.CI)

# Print the cross-validated AUC score
print(paste("Cross-Validated AUC Score:", cv_auc_score.CI))
```

#Infection Contrast Comparisons 

```{r Using A_Temp * Treatment}
#Model used with treatment (combining amplitude and duration) x average temperature
ten.treat.AT.CI <- glm(Infection ~ A_Temp * Treatment, family = binomial, data = CSTC)

# Create custom contrasts for each level of A_Temp
# We want to compare each treatment level to the reference level (Treatment = "0 0")
EMM.CI <- emmeans(ten.treat.AT.CI, ~ A_Temp * Treatment, data = CSTC)
EMM.CI
```

Infinite degrees of freedom due to the statistical output in the model used 

The confidence intervals are likely very wide because of the large standard errors associated with these estimates, which is a common occurrence when you have small sample sizes or sparse data in certain combinations of factors.

```{r Matrices for Contrasts}
#Each row corresponds to one row from EMM.CI

T14.00.CI <- c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T17.00.CI <- c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) 
T20.00.CI <- c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T23.00.CI <- c(0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T14.33.CI <- c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T17.33.CI <- c(0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T20.33.CI <- c(0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0)
T23.33.CI <- c(0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0)
T14.36.CI <- c(0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0)
T17.36.CI <- c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0)
T20.36.CI <- c(0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0)
T23.36.CI <- c(0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0)
T14.63.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0)
T17.63.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0)
T20.63.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0)
T23.63.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0)
T14.66.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0) 
T17.66.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0)
T20.66.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0)
T23.66.CI <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1)
```

```{r Infection contrast comparisons}
contrast(EMM.CI, adjust = "BH", method = list(
  "14 -33 v 00"  = T14.33.CI - T14.00.CI,
  "14 -36 v 00"  = T14.36.CI - T14.00.CI,
  "14 -63 v 00"  = T14.63.CI - T14.00.CI,
  "14 -66 v 00"  = T14.66.CI - T14.00.CI,
  "14 -36 v -33" = T14.36.CI - T14.33.CI,
  "14 -66 v -63" = T14.66.CI - T14.63.CI,
  "14 -63 v -33" = T14.63.CI - T14.33.CI,
  "14 -66 v -36" = T14.66.CI - T14.36.CI,
  
  "17 -33 v 00"  = T17.33.CI - T17.00.CI,
  "17 -36 v 00"  = T17.36.CI - T17.00.CI,
  "17 -63 v 00"  = T17.63.CI - T17.00.CI,
  "17 -66 v 00"  = T17.66.CI - T17.00.CI,
  "17 -36 v -33" = T17.36.CI - T17.33.CI,
  "17 -66 v -63" = T17.66.CI - T17.63.CI,
  "17 -63 v -33" = T17.63.CI - T17.33.CI,
  "17 -66 v -36" = T17.66.CI - T17.36.CI,
  
  "20 -33 v 00"  = T20.33.CI - T20.00.CI,
  "20 -36 v 00"  = T20.36.CI - T20.00.CI,
  "20 -63 v 00"  = T20.63.CI - T20.00.CI,
  "20 -66 v 00"  = T20.66.CI - T20.00.CI,
  "20 -36 v -33" = T20.36.CI - T20.33.CI,
  "20 -66 v -63" = T20.66.CI - T20.63.CI,
  "20 -63 v -33" = T20.63.CI - T20.33.CI,
  "20 -66 v -36" = T20.66.CI - T20.36.CI, 
  
  "23 -33 v 00"  = T23.33.CI - T23.00.CI,
  "23 -36 v 00"  = T23.36.CI - T23.00.CI,
  "23 -63 v 00"  = T23.63.CI - T23.00.CI,
  "23 -66 v 00"  = T23.66.CI - T23.00.CI,
  "23 -36 v -33" = T23.36.CI - T23.33.CI,
  "23 -66 v -63" = T23.66.CI - T23.63.CI,
  "23 -63 v -33" = T23.63.CI - T23.33.CI,
  "23 -66 v -36" = T23.66.CI - T23.36.CI))
```

#Infection Graphs

```{r Ten Infection Percentage and Models}
#making a ggplot plotting by treatment looking at all treatments
TenAvg.CI <- ggplot(CST, aes(x = A_Temp, y = Prop, colour = Treatment)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_smooth(aes(x = A_Temp, y = Prop, colour = Treatment), lty = "dashed",
              method = "glm", method.args = list(family = "binomial"), formula = y ~ poly(x,2),
              se = FALSE, size = 0.6) +
  geom_errorbar(aes(ymin = lower.I, ymax = upper.I), width = 0.2, position = position_dodge(width = 0.5)) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(legend.position = "top", legend.key=element_rect(fill="white"),
        legend.text = element_text(size = 14), legend.title = element_blank()) +
  scale_colour_manual(values = c("#BBBBBB", "#94CBEC", "#5DA899", "#337538", "#2E2585"),
                      labels = c("Constant","-3°C for 3 days", "-3°C for 6 days", "-6°C for 3 days", "-6°C for 6 days")) + 
  theme(axis.text = element_text(size = 14)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  scale_x_continuous(breaks = seq(8, 23, 3)) +
  coord_cartesian(ylim = c(0,1))  # Set the y-axis limits from 0 to 1 as it is proportion data 

TenAvg.CI  

ggsave("AllTenAvg.CI.png", width = 8.5, height = 6)
```

#Infection Interactions Plot

Grouping Points by Duration 

```{r Making Subsets for Interactions (Day 10)}
#Subset the data frames so each has one timing, one temperature, and only cold spell treatments).
TenInt14.D.CI <- subset(binconfCST, A_Temp == "14" & Amplitude %in% c("-3","-6"))
TenInt17.D.CI <- subset(binconfCST, A_Temp == "17" & Amplitude %in% c("-3","-6")) 
TenInt20.D.CI <- subset(binconfCST, A_Temp == "20" & Amplitude %in% c("-3","-6"))
TenInt23.D.CI <- subset(binconfCST, A_Temp == "23" & Amplitude %in% c("-3","-6"))
```

```{r 14 Degree Amplitude vs Duration (Day 10)}
#grouping so amplitude is the x axis, infection is the y axis, and colour grouped by duration
TenInt14.Plot.D.CI <- ggplot(TenInt14.D.CI, aes(x = Amplitude, y = Prop, colour = Duration)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Lower_Conf, ymax = Upper_Conf), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_line(aes(group = Duration), size = 0.75, position = position_dodge(width = 0.3)) + 
  geom_hline(yintercept = 0.95, color = "grey", linetype = "dashed") + # line to show constant infection for the tempertaure
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 1)) 
   
TenInt14.Plot.D.CI
```

```{r 17 Degree Amplitude vs Duration (Day 10)}
TenInt17.Plot.D.CI <- ggplot(TenInt17.D.CI, aes(x = Amplitude, y = Prop, colour = Duration)) + 
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Lower_Conf, ymax = Upper_Conf), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_line(aes(group = Duration), size = 0.75, position = position_dodge(width = 0.3)) + 
  geom_hline(yintercept = 1, color = "grey", linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 1)) 

TenInt17.Plot.D.CI
```

```{r 20 Degree Amplitude vs Duration (Day 10)}
TenInt20.Plot.D.CI <- ggplot(TenInt20.D.CI, aes(x = Amplitude, y = Prop, colour = Duration)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Lower_Conf, ymax = Upper_Conf), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_line(aes(group = Duration), size = 0.75, position = position_dodge(width = 0.3)) + 
  geom_hline(yintercept = 0.92, color = "grey", linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 1)) 

TenInt20.Plot.D.CI
```

```{r 23 Degree Amplitude vs Duration (Day 10)}
TenInt23.Plot.D.CI <- ggplot(TenInt23.D.CI, aes(x = Amplitude, y = Prop, colour = Duration)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Lower_Conf, ymax = Upper_Conf), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_line(aes(group = Duration), size = 0.75, position = position_dodge(width = 0.3)) + 
  geom_hline(yintercept = 0.71, color = "grey", linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_y_continuous(position = "right") +
  theme(axis.text = element_text(size = 14)) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = c(0.5,0.2), legend.key=element_rect(fill="white"),
        legend.text = element_text(size = 12), legend.title = element_text(size = 12)) +
  scale_colour_manual("CS Duration", values = c("#FF6666","#33CCCC"), 
                      labels = c("3 Days", "6 Days")) 

TenInt23.Plot.D.CI
```

#Exposed Model Test

```{r Exposed zero inflated model}
# Fit a zero-inflated negative binomial model
zinb_model.CE <- zeroinfl(Exposed ~ poly(Real_Avg_Cent, 3) * Treatment | poly(Real_Avg_Cent, 3) * Treatment, data = CST, dist = 'negbin')

# Summary of the model
summary(zinb_model.CE).CE
```

```{r Exposed GLM}
ten.nb.CE <- glm.nb(Exposed ~ poly(Real_Avg_Cent, 3) * Amplitude * Duration, data = CST)
summary(ten.nb.CE)

ten.anova.CE <- anova(ten.nb.CE, test = "Chi")
ten.anova.CE
```

```{r Burden GLM}
ten.nb.CB <- glm.nb(Spores ~ poly(Real_Avg_Cent, 3) * Amplitude * Duration, data = CST)

summary(ten.nb.CB)

ten.anova.CB <- anova(ten.nb.CB, test = "Chi")
ten.anova.CB
```

#Burden Model

```{r Centred Cubic NB Burden Model}
#Lowest AIC found a cubic fit was most suitable for the data as it is nonlinear
#use data set with NAs removed as they were uninfected and so never had a spore burden
ten.nb.CB <- glm.nb(Spores ~ poly(Real_Avg_Cent, 3) * Amplitude * Duration, data = CST.NA)
summary(ten.nb.CB)

ten.anova.CB <- anova(ten.nb.CB, test = "Chi")
ten.anova.CB
```

Again, colinearity deems it inaccurate model and from the basis of this model an elastic regression is created 

```{r Negative Binomial Elastic Regression}
# Prepare the data
x.CB <- model.matrix(Spores ~ poly(Real_Avg_Cent, 3) * Amplitude * Duration, data = CST.NA)
y.CB <- CST.NA$Spores
```

```{r Finding optimal alpha}
#different way to find the most suitable alpha but still involves modelling and finding the lowest AIC out of 1, 0.5, 0
#theta needs to be manually specified in glmnet so we used the theta present in the GLM 
fit_lasso.CB <- glmnet(x.CB, y.CB, family = negative.binomial(theta = 1.2208), alpha = 1)
fit_mix.CB <- glmnet(x.CB, y.CB, family = negative.binomial(theta = 1.2208), alpha = 0.5)
fit_ridge.CB<- glmnet(x.CB, y.CB, family = negative.binomial(theta = 1.2208), alpha = 0)

# Calculate deviance for each model
dev_lasso.CB <- deviance(fit_lasso.CB)
dev_mix.CB <- deviance(fit_mix.CB)
dev_ridge.CB <- deviance(fit_ridge.CB)

# Calculate number of non-zero coefficients for each model
num_nonzero_lasso.CB <- sum(coef(fit_lasso.CB, s = 0) != 0)
num_nonzero_mix.CB <- sum(coef(fit_mix.CB, s = 0) != 0)
num_nonzero_ridge.CB <- sum(coef(fit_ridge.CB, s = 0) != 0)

# Calculate AIC and BIC
n.CB <- length(y.CB)
aic_lasso.CB <- n.CB * log(dev_lasso.CB / n.CB) + 2 * num_nonzero_lasso.CB
aic_mix.CB <- n.CB * log(dev_mix.CB / n.CB) + 2 * num_nonzero_mix.CB
aic_ridge.CB <- n.CB * log(dev_ridge.CB / n.CB) + 2 * num_nonzero_ridge.CB

bic_lasso.CB <- n.CB * log(dev_lasso.CB / n.CB) + log(n.CB) * num_nonzero_lasso.CB
bic_mix.CB <- n.CB * log(dev_mix.CB / n.CB) + log(n.CB) * num_nonzero_mix.CB
bic_ridge.CB <- n.CB * log(dev_ridge.CB / n.CB) + log(n.CB) * num_nonzero_ridge.CB

# Compare AIC or BIC values to choose the best model
if (aic_lasso.CB < aic_mix.CB & aic_lasso.CB < aic_ridge.CB) {
  best_alpha_aic.CB = 1
} else if (aic_mix.CB < aic_ridge.CB) {
  best_alpha_aic.CB = 0.5
} else {
  best_alpha_aic.CB = 0
}

if (bic_lasso.CB < bic_mix.CB & bic_lasso.CB < bic_ridge.CB) {
  best_alpha_bic.CB = 1
} else if (bic_mix.CB < bic_ridge.CB) {
  best_alpha_bic.CB = 0.5
} else {
  best_alpha_bic.CB = 0
}

# Print the best alpha values based on AIC and BIC
print(paste("Best alpha based on AIC:", best_alpha_aic.CB))
print(paste("Best alpha based on BIC:", best_alpha_bic.CB))
```

```{r Initial Model fit}
# Fit the elastic net model using the theta from the glm.nb 
elfit.nb.CB <- glmnet(x.CB, y.CB, family = negative.binomial(theta = 1.2208), alpha = 0.5)
```

```{r Finding the optimal lambda}
# Calculate AIC for each lambda
aic_values <- sapply(elfit.nb.CB$lambda, function(lambda) {
  fitted_values <- predict(elfit.nb.CB, newx = x.CB, s = lambda, type = "response")
  logLikelihood <- sum(dnbinom(y.CB, mu = fitted_values, size = 1.2208, log = TRUE))
  k <- sum(coef(elfit.nb.CB, s = lambda, exact = TRUE) != 0)
  -2 * logLikelihood + 2 * k
})

# Calculate BIC for each lambda
n <- length(y.CB)
bic_values <- sapply(elfit.nb.CB$lambda, function(lambda) {
  fitted_values <- predict(elfit.nb.CB, newx = x.CB, s = lambda, type = "response")
  logLikelihood <- sum(dnbinom(y.CB, mu = fitted_values, size = 1.2208, log = TRUE))
  k <- sum(coef(elfit.nb.CB, s = lambda, exact = TRUE) != 0)
  -2 * logLikelihood + log(n) * k
})

# Print AIC and BIC values
AIC.BIC.CB <- data.frame(lambda = elfit.nb.CB$lambda, AIC = aic_values, BIC = bic_values)
```

```{r Optimal Lambda Elastic Regression}
lambda.CB <- elfit.nb.CB$lambda
#set the lambda as the lowest AIC 
lambda.AIC.CB <- 0.071067144

#final regression with the lowest AIC valued lambda 
elfit.aic.CB <- glmnet(x.CB, y.CB, family = negative.binomial(theta = 1.2208), alpha = 0.5, lambda = lambda.AIC.CB)
```

```{r Bootstrap significance testing for burden}
# Number of bootstrap iterations
num_bootstraps.CB <- 25000

# Empty list to store bootstrap coefficients
bootstrap_coefs_list.CB <- vector("list", length = num_bootstraps.CB)

# Set the seed for reproducibility
set.seed(123)

# Perform bootstrapping
for (i in 1:num_bootstraps.CB) {
  # Sample with replacement from the data
  sample_indices.CB <- sample(nrow(x.CB), replace = TRUE)
  x_sampled.CB <- x.CB[sample_indices.CB, ]
  y_sampled.CB <- y.CB[sample_indices.CB]
  
  # Fit the elastic net model on the sampled data
  model_fit.CB <- glmnet(x_sampled.CB, y_sampled.CB, family = negative.binomial(theta = 1.2208), alpha = 0.5, lambda = lambda.AIC.CB)
  
  # Store the coefficients in the list
  bootstrap_coefs_list.CB[[i]] <- as.vector(coef(model_fit.CB, s = lambda.AIC.CB))
}

# Convert the list of bootstrap coefficients to a matrix
bootstrap_coefs.CB <- do.call(rbind, bootstrap_coefs_list.CB)

# Calculate the 95% confidence intervals for each coefficient
lower_ci.CB <- apply(bootstrap_coefs.CB, 2, function(coef_col) quantile(coef_col, 0.025)) #2.5th percentile
upper_ci.CB <- apply(bootstrap_coefs.CB, 2, function(coef_col) quantile(coef_col, 0.975)) #97.5th percentile 

# Convert the coefficients to a matrix
coef_matrix.CB <- as.matrix(coef(elfit.aic.CB, s = lambda.AIC.CB))

# Combine the results into a data frame for easy interpretation
coef_summary.CB <- data.frame(Coefficient = colnames(coef_matrix.CB),
                              Estimate = coef_matrix.CB,
                              Lower.CB = lower_ci.CB,
                              Upper.CB = upper_ci.CB)

# Print the summary
print(coef_summary.CB)
```

```{r Feature Importance for Burden}
# Extract the coefficients from the final model
coef_matrix.CB <- as.matrix(coef(elfit.aic.CB, s = lambda.AIC.CB))

# Get the absolute values of coefficients to identify feature importance
feature_importance.CB <- abs(coef_matrix.CB[-1,])

# Sort the features by importance (highest to lowest)
sorted_features.CB <- feature_importance.CB[order(-feature_importance.CB)]

# Get the corresponding feature names
sorted_feature_names.CB <- colnames(coef_matrix.CB)[-1][order(-feature_importance.CB)]

# Combine the feature names with the sorted feature importance values
sorted_feature_importance_table.CB <- data.frame(Feature = sorted_feature_names.CB,
                                                 Importance = sorted_features.CB)

# Print the sorted feature importance table
print(sorted_feature_importance_table.CB)
```

```{r Predicted vs Actual Data}
# Assuming 'predictions' contains the predicted values and 'y.CB' contains the actual values
predict.CB <- predict(elfit.nb.CB, newx = x.CB, s = lambda.AIC.CB)

# Create a data frame for the plot
predict.plot.CB<- data.frame(Actual = y.CB, Predicted = predict.CB)

# Create the prediction plot using ggplot2
ggplot(predict.plot.CB, aes(x = Actual, y = s1)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "Prediction Plot",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal()
```

#Burden Contrast Comparisons

```{r Using A_Temp * Treatment}
#Model used again using treatment (combines amplitude and duration)
ten.treat.AT.CB <- glm.nb(Spores ~ A_Temp * Treatment, data = CSTC)

EMM.CB <- emmeans(ten.treat.AT.CB, ~ A_Temp * Treatment, data = CSTC)
EMM.CB
```

```{r Matrices for Contrasts}
#Each line corresponds to a line in EMM.CB
T14.00.CB <- c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T17.00.CB <- c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) 
T20.00.CB <- c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T23.00.CB <- c(0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T14.33.CB <- c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T17.33.CB <- c(0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
T20.33.CB <- c(0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0)
T23.33.CB <- c(0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0)
T14.36.CB <- c(0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0)
T17.36.CB <- c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0)
T20.36.CB <- c(0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0)
T23.36.CB <- c(0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0)
T14.63.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0)
T17.63.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0)
T20.63.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0)
T23.63.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0)
T14.66.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0) 
T17.66.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0)
T20.66.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0)
T23.66.CB <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1)
```

```{r Burden Custom Contrasts}
#Comparing lines 
contrasts.CB <- contrast(EMM.CB, adjust = "BH", method = list(
  "14 -33 v 00"  = T14.33.CB - T14.00.CB,
  "14 -36 v 00"  = T14.36.CB - T14.00.CB,
  "14 -63 v 00"  = T14.63.CB - T14.00.CB,
  "14 -66 v 00"  = T14.66.CB - T14.00.CB,
  "14 -36 v -33" = T14.36.CB - T14.33.CB,
  "14 -66 v -63" = T14.66.CB - T14.63.CB,
  "14 -63 v -33" = T14.63.CB - T14.33.CB,
  "14 -66 v -36" = T14.66.CB - T14.36.CB,
  
  "17 -33 v 00"  = T17.33.CB - T17.00.CB,
  "17 -36 v 00"  = T17.36.CB - T17.00.CB,
  "17 -63 v 00"  = T17.63.CB - T17.00.CB,
  "17 -66 v 00"  = T17.66.CB - T17.00.CB,
  "17 -36 v -33" = T17.36.CB - T17.33.CB,
  "17 -66 v -63" = T17.66.CB - T17.63.CB,
  "17 -63 v -33" = T17.63.CB - T17.33.CB,
  "17 -66 v -36" = T17.66.CB - T17.36.CB,
  
  "20 -33 v 00"  = T20.33.CB - T20.00.CB,
  "20 -36 v 00"  = T20.36.CB - T20.00.CB,
  "20 -63 v 00"  = T20.63.CB - T20.00.CB,
  "20 -66 v 00"  = T20.66.CB - T20.00.CB,
  "20 -36 v -33" = T20.36.CB - T20.33.CB,
  "20 -66 v -63" = T20.66.CB - T20.63.CB,
  "20 -63 v -33" = T20.63.CB - T20.33.CB,
  "20 -66 v -36" = T20.66.CB - T20.36.CB, 
  
  "23 -33 v 00"  = T23.33.CB - T23.00.CB,
  "23 -36 v 00"  = T23.36.CB - T23.00.CB,
  "23 -63 v 00"  = T23.63.CB - T23.00.CB,
  "23 -66 v 00"  = T23.66.CB - T23.00.CB,
  "23 -36 v -33" = T23.36.CB - T23.33.CB,
  "23 -66 v -63" = T23.66.CB - T23.63.CB,
  "23 -63 v -33" = T23.63.CB - T23.33.CB,
  "23 -66 v -36" = T23.66.CB - T23.36.CB))

summary(contrasts.CB)
```

#Burden Graphs

```{r Ten Average Burden and Model}
#making a ggplot plotting by treatment looking at all treatments
TenAvg.CB <- ggplot(CST, aes(x = A_Temp, y = Spores, colour = Treatment)) +
  geom_smooth(aes(x = A_Temp, y = Spores, colour = Treatment),
              method = "glm.nb", formula = y ~ poly(x, 3),
              se = FALSE, size = 0.5, lty = "dashed") +
  stat_summary(fun = "mean", geom = "point", size = 2.5,
               position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = mean - se.B, ymax =  mean + se.B), width = 0.2, position = position_dodge(width = 0.5)) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  theme(legend.title = element_text(face = "bold")) +
  theme(legend.position = "top", legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 14), legend.title = element_blank()) +
  scale_colour_manual(values = c("#BBBBBB", "#94CBEC", "#5DA899", "#337538", "#2E2585"),
                      labels = c("Constant","-3°C for 3 days", "-3°C for 6 days", "-6°C for 3 days", "-6°C for 6 days")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  scale_x_continuous(breaks = seq(8, 23, 3)) +
  coord_cartesian(ylim = c(0,700))

TenAvg.CB

ggsave("AllTenAvg.CB.png", width = 8.5, height = 6)

# old colours = "#999999", "#57B4E9", "#0D72B2", "#CC79A9", "#70135C"
```

#Burden Interaction Plots

```{r Making Subsets for Interactions (Day 10)}
#Subset the data frames so each has one timing, one temperature, and only cold spell treatments).
TenInt14.D.CB <- subset(CST, A_Temp == "14" & Amplitude %in% c("-3","-6"))
TenInt17.D.CB <- subset(CST, A_Temp == "17" & Amplitude %in% c("-3","-6")) 
TenInt20.D.CB <- subset(CST, A_Temp == "20" & Amplitude %in% c("-3","-6"))
TenInt23.D.CB <- subset(CST, A_Temp == "23" & Amplitude %in% c("-3","-6"))
```

```{r 14 Degree Amplitude vs Duration (Day 10)}
TenInt14.Plot.D.CB <- ggplot(TenInt14.D.CB, aes(x = Amplitude, y = Spores, colour = Duration)) +
  stat_summary(fun = "mean", geom = "point", size = 2, position = position_dodge(width = 0.3)) +
  stat_summary(fun = "mean", geom = "line", aes(group = Duration), size = 0.75, position = position_dodge(width = 0.3)) + 
  geom_errorbar(aes(ymin = mean - se.B, ymax =  mean + se.B), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept = 98, color = "grey", linetype = "dashed") + # Line for constant treatment
  geom_hline(yintercept = 78, color = "#ECECEC") + #lower SE (SE = 20)
  geom_hline(yintercept = 118, color = "#ECECEC") + #Upper SE (SE = 20)
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 750)) +
  theme(legend.position = c(0.5,0.8), legend.key=element_rect(fill="white"),
        legend.text = element_text(size = 12), legend.title = element_text(size = 12)) 

TenInt14.Plot.D.CB
ggsave("TenInt14.Plot.D.CB.png", width = 2.5, height = 3.5)
```
```{r 17 Degree Amplitude vs Duration (Day 10)}
TenInt17.Plot.D.CB <- ggplot(TenInt17.D.CB, aes(x = Amplitude, y = Spores, colour = Duration)) + 
  stat_summary(fun = "mean", geom = "point", size = 2,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = "mean", geom = "line", aes(group = Duration), size = 0.75,
               position = position_dodge(width = 0.3)) + 
  geom_errorbar(aes(ymin = mean - se.B, ymax =  mean + se.B), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept = 427, color = "grey", linetype = "dashed") +
  geom_hline(yintercept = 386, color = "#ECECEC") + #lower SE (SE = 41)
  geom_hline(yintercept = 468, color = "#ECECEC") + #Upper SE (SE = 41)
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  scale_y_continuous(position = "right") +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 750)) 

TenInt17.Plot.D.CB
ggsave("TenInt17.Plot.D.CB.png", width = 2.5, height = 3.5)
```
```{r 20 Degree Amplitude vs Duration (Day 10)}
TenInt20.Plot.D.CB <- ggplot(TenInt20.D.CB, aes(x = Amplitude, y = Spores, colour = Duration)) +
  stat_summary(fun = "mean", geom = "point", size = 2,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = "mean", geom = "line", aes(group = Duration), size = 0.75,
               position = position_dodge(width = 0.3)) + 
  geom_errorbar(aes(ymin = mean - se.B, ymax =  mean + se.B), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept = 486, color = "grey", linetype = "dashed") +
  geom_hline(yintercept = 425, color = "#ECECEC") + #lower SE (SE = 61)
  geom_hline(yintercept = 547, color = "#ECECEC") + #Upper SE (SE = 61)
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  scale_y_continuous(position = "right") +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 750)) 

TenInt20.Plot.D.CB
ggsave("TenInt20.Plot.D.CB.png", width = 2.5, height = 3.5)
```
```{r 23 Degree Amplitude vs Duration (Day 10)}
TenInt23.Plot.D.CB <- ggplot(TenInt23.D.CB, aes(x = Amplitude, y = Spores, colour = Duration)) +
  stat_summary(fun = "mean", geom = "point", size = 2,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = "mean", geom = "line", aes(group = Duration), size = 0.75,
               position = position_dodge(width = 0.3)) + 
  geom_errorbar(aes(ymin = mean - se.B, ymax =  mean + se.B), width = 0.2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept = 89, color = "grey", linetype = "dashed") +
  geom_hline(yintercept = 64, color = "#ECECEC") + #lower SE (SE = 25)
  geom_hline(yintercept = 114, color = "#ECECEC") + #Upper SE (SE = 25)
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  scale_y_continuous(position = "right") +
  guides(color = "none") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  coord_cartesian(ylim = c(0, 750)) 

TenInt23.Plot.D.CB
ggsave("TenInt23.Plot.D.CB.png", width = 2.5, height = 3.5)
```
